---
layout: page
title: "Playing around with HuffPollster data"
author: "Olivia Brode-Roger"
date: "February 16, 2016"
permalink : /polls/
---

```{r setup, include=FALSE}
library(ggplot2)
library(reshape2)
library(pollstR)
```

```{r, echo=FALSE, warning=FALSE, fig.width=17, fig.height=10}
d_polls <- pollstr_polls(topic = "2016-president-dem-primary", state = "US", after="2016-01-01", max_pages=500)
d_candidates <- c("Clinton", "Sanders")
d_questions <- subset(d_polls[["questions"]], question == "2016 National Democratic Primary" & choice %in% d_candidates & value > 0)
d_polldata <- merge(d_polls$polls, d_questions, by = "id")

p <- ggplot(d_polldata, aes(x = end_date, y = log(value/(100-value)), color = choice)) + xlim(as.Date(c("2016-01-01", "2016-06-15")))
p <- p + geom_point(alpha = 0.2)
p <- p + geom_smooth(span = 0.5, method="loess")

targets <- read.csv("2016-dem-primary-targets-manual.csv")
results <- read.csv("2016-dem-primary.csv")

targets$state_abb <- state.abb[match(targets$state, state.name)]
targets$clinton_result <- log(results$Clinton_popular_vote/(100-results$Clinton_popular_vote))
targets$sanders_result <- log(results$Sanders_popular_vote/(100-results$Sanders_popular_vote))
targets$date <- as.Date(results$Date)
state_clinton_targets <- melt(targets, id.vars = c("state", "state_abb", "date", "sanders_result", "clinton_result"))

p_states <- p +
  geom_point(data = state_clinton_targets, mapping=aes(x=date, y = clinton_result-value, color = variable), size=3) +
  geom_line(data = state_clinton_targets, mapping=aes(x=date, y = clinton_result-value, color = variable), alpha = 0.8) +
  geom_point(data = state_clinton_targets, mapping=aes(x=date, y = sanders_result+value, color = variable), size=3) +
  geom_line(data = state_clinton_targets, mapping=aes(x=date, y = sanders_result+value, color = variable), alpha = 0.8) +
  xlab("Date") + ylab("Polls and result (log odds ratio)") +
  theme(text = element_text(size=20))
p_states
```

This chart seeks to satisfy some of my more particular desires in charts, as well as add a bit of information.
I am only using the democratic primary race, due to a target analysis that is *significantly* easier to run in a 2-person race.

- better moving average (with std error!)
- plotting empty space for the future
- I don't like percentages much, I significantly prefer [log odds-ratio]({% post_url 2016-02-13-why-I-prefer-log-odds-ratio %})
- plotting the state results after using [targets to correct them for bias]({site_url}/targets).

This doesn't tell us much (yet!), but these new points probably tell us something about how the rest of the country is going to vote, assuming nothing changes.