---
layout: page
title: "Target tracking"
author: "Olivia Brode-Roger"
date: "February 16, 2016"
permalink : /targets/
---

```{r setup, include=FALSE}
require(ggplot2)
require(dplyr)
require(reshape2)
```

States above the diagonal are where Clinton exceeded targets, below Sanders is exceeding target, more in my (first!) [blog post]({% post_url 2016-02-10-targets-v-results %})
Arbitrarely, positive indicates "Clinton-ness", 0 is neutral and negative is "Sanders-ness".

```{r, echo=FALSE, fig.width=17, fig.height=11, warning=FALSE}
targets <- read.csv("2016-dem-primary-targets-manual.csv")
results <- read.csv("2016-dem-primary.csv")

results <- mutate(results, state_abb = state.abb[match(state, state.name)]) %>%
  mutate(result = log(Clinton_popular_vote/Sanders_popular_vote)) %>%
  select(state, state_abb, result)
targets <- merge(targets, results, by.x = "state")
state_targets <- melt(targets, id.vars = c("state", "state_abb", "result"))

ggplot(state_targets, aes(x = value, y = result, color=variable) ) +
  geom_text( aes(label=state_abb) ) +
  labs(title = "Various Targets v Results", x = "Targets", y = "Results", size = "Total delegates") +
  geom_abline(intercept = 0, slope=1, alpha = 0.5) +
  theme(text = element_text(size=20)) +
  #xlim(-1.2,1.2) + ylim(-1.2,1.2) +
  coord_equal() +
  facet_wrap(~ variable) +
  theme(legend.position = "none")
```

Because the distance from the diagonal is what we care about, here's a change of coordinates:
```{r, echo=FALSE, fig.width=17, fig.height=11, warning=FALSE}
ggplot(state_targets, aes(x = value, y = result-value) ) +
  geom_text( aes(label=state_abb) ) +
  geom_smooth(method="lm", alpha = 0.2) +
  labs(title = "Various Targets v Diff", x = "Targets", y = "Bias-corrected result") +
  geom_abline(intercept = 0, slope=0, alpha = 0.5) +
  theme(text = element_text(size=20)) +
  coord_equal() +
  facet_wrap(~ variable) +
  theme(legend.position = "none")
```
The targets used are:

- [Cook Political Report, Jan 21](http://cookpolitical.com/story/9179).
- [Cook Political Report, Feb 12](http://cookpolitical.com/story/9258).
- [my first attempt]({% post_url 2016-02-15-targets---FiveThirtyEight-edition %}) (`white*2012 Democratic vote share`)
- [FiveThirtyEight's targets](http://fivethirtyeight.com/features/bernie-sanderss-path-to-the-nomination/)
