---
layout: page
title: "Target tracking"
author: "Olivia Brode-Roger"
date: "February 16, 2016"
permalink : /targets/
---

```{r setup, include=FALSE}
require(ggplot2)
require(reshape2)
```

An explanation for these plot can be found with [my first blog post]({% post_url 2016-02-10-targets-v-results %}).
Essentially, states above the diagonal are where Clinton exceeded targets, below Sanders is exceeding target.
Arbitrarely, positive indicates "Clinton-ness", 0 is neutral and negative is "Sanders-ness".

```{r, echo=FALSE, fig.width=11, warning=FALSE}
targets <- read.csv("2016-dem-primary-targets.csv")
results <- read.csv("2016-dem-primary.csv")

state_targets <- data.frame(state = state.abb[match(targets$state, state.name)],
                     CPR_Jan21 = targets$CPR_Clinton_target,
                     Olivia = targets$my_Clinton_target,
                     result = log(results$Clinton_popular_vote/results$Sanders_popular_vote))
state_targets <- melt(state_targets, id.vars = c("state", "result"))

plot <- ggplot(state_targets, aes(x = value, y = result, color=variable) ) +
  geom_text( aes(label=state) ) +
  #scale_size_continuous(range = c(2,7), limits = c(14,475)) +
  labs(title = "Various Targets v Results", x = "Targets", y = "Results", size = "Total delegates") +
  geom_abline(intercept = 0, slope=1, alpha = 0.5) +
  xlim(-1,1) + ylim(-1,1)
plot
```

The targets used are:

- [Cook Political Report, Jan 21](http://cookpolitical.com/story/9179).
- [my first attempt]({% post_url 2016-02-15-targets---FiveThirtyEight-edition %}) (`white*2012 Democratic vote share`)
- more to come soon!
